### 2015-06-18 chapters/ipa_meets_unicode.tex
added text on tie-bar; updated gitignore

Finally, the tie bar to indicate affricates, doubly articulated consonants or
diphthongs can be placed either below or on top of the base characters, e.g.
<\charis{t͡s}> or <\charis{t͜s}>. For strict IPA encoding we propose to standardize on the tie bar
above the base characters, using Unicode \textsc{combining double inverted
breve} at \uni{0361}. Note that this Unicode character is placed between the two 
base characters to be combined. In principle, it is possible to combine more 
than two base characters by repeating the tie bar, like in <a͡o͡u>. If really
necessary, we consider this possible, even though the rendering will never look 
good.

>>>>

Finally, the tie bar to indicate affricates, doubly articulated consonants or
diphthongs can be placed above or below the base characters, e.g.
<\charis{t͡s}> or <\charis{t͜s}>. IPA allows both the \textsc{combining double inverted
breve} at \uni{0361} and the \textsc{combining double breve below} at \uni{035C}. The  
choice between the two symbols is purely for legible rendering; there is no difference 
in semantics between the two symbols. In fact, rendering is such a problematic issue 
for tie bars in general, that many linguists simply do not use them. Further, tie bars 
are a special type of character in the sense that they bind two graphemes (singleton base characters 
or tailored grapheme clusters) together. In Section \ref{pitfall-different-notions-of-diacritics}, 
we highlight the division of linguistic diacritics 
into two categories in the Unicode Standard: Spacing Modifier Letters and Combining 
Diacritical Marks.\footnote{The former contain characters like aspiration < ʰ > which is user 
specified (before or after the base character) and the latter which combines 
with the base character.} The tie-bar is somewhere in between these two character 
classes, i.e. when the combining breve (above or below) is told to combine within 
a sequence of Unicode characters, it will bind to the character immediately 
preceding it -- the so-called base character -- regardless if that base 
character is a Spacing Modifier Letter \textit{or} Combining Diacritical Mark. \footnote{In some cases 
conversion of a string sequence into Unicode Normalization Form C alleviates 
the problem by composing sequences of characters like <e> + < ́ > into single character < é > 
(see however diacritics like the rhotic hook in ).} In principle, 
it is possible to combine more than two base characters by repeating the tie bar, 
like in <a͡o͡u>. If really necessary, we consider this possible, even though the 
rendering will never look good.

The tie-bar is a convenient diacritic for visually tokenizing input strings into chunks of 
phonetically salient groups and both the above and below tie-bar were introduced in electronic 
text to deal with difficult to render character sequences. The tie-bar is unnecessary for 
computational tokenization given the design of orthography profiles, described in 
Section \ref{usecase-tokenization}. However, to adhere to strict IPA encoding users 
may use either tie bar, or none whatsoever.

\begin{comment} 
Finally, the tie bar to indicate affricates, doubly articulated consonants or
diphthongs can be placed either below or on top of the base characters, e.g.
<\charis{t͡s}> or <\charis{t͜s}>. For strict IPA encoding we propose to standardize on the tie bar
above the base characters, using Unicode \textsc{combining double inverted
breve} at \uni{0361}. Note that this Unicode character is placed between the two 
base characters to be combined. In principle, it is possible to combine more 
than two base characters by repeating the tie bar, like in <a͡o͡u>. If really
necessary, we consider this possible, even though the rendering will never look 
good.

For strict IPA encoding we propose to standardize on the tie bar
above the base characters, using Unicode \textsc{combining double inverted
breve} at \uni{0361}. Note that this Unicode character is placed between the two 
base characters to be combined. In principle, it is possible to combine more 
than two base characters by repeating the tie bar, like in <a͡o͡u>. If really
necessary, we consider this possible, even though the rendering will never look 
good.\end{comment}

### 2015-06-18 chapters/use_cases.tex
added text on tie-bar; updated gitignore

We now present several use cases that have motivated the development of
orthography profiles. These include: 
\begin{itemize}
	\item tokenization and error checking
	\item normalization of orthographic systems to attain interoperability across different source documents, 
	\item cognate identification for detecting the similarity between words from different languages 
\end{itemize}

@@

- Unicode character tokenization
- grapheme tokenization
- tailored grapheme tokenization

@@

\citet{Moran2012;MoranMcCloy2014} mark tones as singletons with Space Modifier
Letters, e.g. <˦> for a phonemic high tone, instead of accent diacritics,
alleviating potential conflicts. Building on these works,

>>>>

We now present several use cases that have motivated the development of
orthography profiles. These include:

\begin{itemize}
	\item tokenization and error checking
	\item normalization of orthographic systems to attain interoperability across different source documents, 
	\item cognate identification for detecting the similarity between words from different languages 
	\item ...
\end{itemize}

@@

\begin{itemize}
	\item Unicode character tokenization
	\item grapheme tokenization
	\item tailored grapheme tokenization
\end{itemize}

@@

\citet{Moran2012;MoranMcCloy2014} mark tones as singletons with Space Modifier
Letters, e.g. <˦> for a phonemic high tone, instead of accent diacritics,
alleviating potential conflicts.

\citet{Moran2012;MoranMcCloy2014} represent clicks as a combination of anterior and 
posterior articulations, where the set < k q ɡ ɢ ŋ ɴ > denote place of articulation 
and phonation type (voiced, voiceless, nasal). The anterior articulation click characters 
are < ʘ ǀ ǁ ǃ ǂ > with the addition of non-standard IPA < ‼ > \textsc{DOUBLE EXCLAMATION MARK)} 
\uni{203C} to represent the retroflex click. Note that the (post)alveolar click < ǃ > at \uni{01C3} 
is called incorrectly \textsc{LATIN LETTER RETROFLEX CLICK} (it is also a different character than 
the keyboard < ! > \textsc{EXCLAMATION MARK} at \uni{0021}). All sources that contain clicks in 
PHOIBLE are standardized to contain both the anterior and posterior characters and diacritics 
are added to the posterior glyph, e.g.\ < ŋ̊ǃ >.\footnote{http://phoible.github.io/conventions/}

Building on these works...
